---
title: "DE genes expression - 6OHDA and mitopark models"
author: "Yuvarani Masarapu"
date: "2023-09-05"
output:
  html_document:
    self_contained: true
    number_sections: true
    highlight: tango
    df_print: paged
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r eval=FALSE}
#rmarkdown::render(input = "/Users/yuvarani.masarapu/Documents/parkinsons_project/mitopark-model/DE_genes_comparison.Rmd", output_file = "/Users/yuvarani.masarapu/Documents/parkinsons_project/mitopark-model/expression_plots_top-DEGs.html")
```

Custom function to read a excel file with multiple sheets

```{r}
library(readxl)
read_excel_file <- function(fname){
  
  sheets_names <- readxl::excel_sheets(fname)
  tibble <- lapply(sheets_names, function(x){
    readxl::read_excel(path = fname, sheet = x)
  })
  
  #names(tibble) <- sheets_names
}
```

# Mitopark data (.rds file)
```{r eval=FALSE}
mitopark <- readRDS(file = "rna.combined.clusters.rds")
```

Read the DEGs excel file via custom function
```{r eval=FALSE}
path <- "/Users/yuvarani.masarapu/Documents/parkinsons_project/mitopark-model/mitopark-KO_wt_DEGs-cluster_2023-09-05.xlsx"
deg_mitopark <- read_excel_file(path)

names(deg_mitopark) <- c(0:(length(deg_mitopark) - 1))
```

```{r umap_clusters2, fig.height=8, fig.width=9}
cl.colors <-  c("#F18F01", "#878787", "#FCCDE5", "#048BA8", "#2E4057", "#99C24D", "#B95F89", "#5F070C" ,  "#DFC27D", "#DB6C79", "#B191FF", "#157A6E", "#EB4511", "#73683B", "#97C8EB","#C51B7D", "#BA9763", "#31081F", "#52D1DC", "#700353", "#F5D3C8", "#725752", "#D8315B", "#6B4E71", "#8CBFCF" , "#C2B2B4", "#690375" ,"#EDE5A6", "#52B788", "#EE1B26", "#F2F230", "#91F291", "#845A6D")
#pie(rep(1,length(cl.colors)), col = cl.colors)

DimPlot(object = mitopark, group.by = "seurat_clusters", reduction = "umap_after_harmony", pt.size = 0.6, label = T, label.size = 6, cols = cl.colors) + ggtitle("Clusters in mitopark snRNAseq dataset")
```

## Expression plots of top 5 candidate DEGs per cluster (p_val_adj < 0.01 and reordered by average logFC values)

* Less than 5 genes mean that the other DEGs in that cluster values less than p_val_adj of 0.01.

```{r plot_mitopark, warning=FALSE, message=FALSE, results='asis', fig.width=7, fig.height=7}
for(cluster in names(deg_mitopark)){
  cat(paste("\n\n### Cluster ",cluster, "\n\n"))
  
  top_genes <- deg_mitopark[[as.integer(cluster) + 1]] %>%
    filter(p_val_adj < 0.01) %>%
    arrange(avg_log2FC) %>%
    tail(n = 5) %>%
    select(gene) 
  top_genes <- top_genes[,1]
  
  if(length(rownames(top_genes))>0){
    
    for(gene in top_genes){
      DefaultAssay(mitopark) <- "SCT"
      print(FeaturePlot(object = mitopark, features = gene, reduction = "umap_after_harmony", slot = "counts"))
    }
  }
}
```

# 6OHDA data (.rds file)
```{r eval=FALSE}
OHDA <- readRDS(file = "/Users/yuvarani.masarapu/Documents/parkinsons_project/sn_data/test_parkinsons_sn/rna.combined.clusters.rds")
#clusters.list <- readRDS("~/Documents/parkinsons_project/sn_data/test_parkinsons_sn/rna.combined_split-clusters.rds")
```

Read the DEGs excel file via custom function
```{r eval=FALSE}
path <- "/Users/yuvarani.masarapu/Documents/parkinsons_project/sn_data/test_parkinsons_sn/6OHDA-vs-control_DEGs-cluster_2023-07-10.xlsx"
deg_6OHDA <- read_excel_file(path)

names(deg_6OHDA) <- c(0:(length(deg_6OHDA) - 1))
```

```{r fig.height=8, fig.width=9}
cl.colors <-  c("#F18F01", "#878787", "#FCCDE5", "#048BA8", "#2E4057", "#99C24D", "#B95F89", "#5F070C" ,  "#DFC27D", "#DB6C79", "#B191FF", "#157A6E", "#EB4511", "#73683B", "#97C8EB","#C51B7D", "#BA9763", "#31081F", "#52D1DC", "#700353", "#F5D3C8", "#725752", "#D8315B", "#6B4E71", "#8CBFCF" , "#C2B2B4", "#690375" ,"#EDE5A6", "#52B788", "#EE1B26", "#F2F230", "#91F291", "#845A6D")
#pie(rep(1,length(cl.colors)), col = cl.colors)

DimPlot(object = OHDA, group.by = "seurat_clusters", reduction = "umap_after_harmony", pt.size = 0.6, label = T, label.size = 6, cols = cl.colors) + ggtitle("Clusters for 6OHDA snRNAseq dataset")
```

## Expression plots of top 5 candidate DEGs per cluster (p_val_adj < 0.01 and reordered by average logFC values)

* Less than 5 genes mean that all other DEGs in that cluster are not significant and have values less than p_val_adj of 0.01.

```{r plot_6OHDA, warning=FALSE, message=FALSE, results='asis', fig.width=7, fig.height=7}
for(cluster in names(deg_6OHDA)){
  cat(paste("\n\n### Cluster ",cluster, "\n\n"))
  
  top_genes <- deg_6OHDA[[as.integer(cluster) + 1]] %>%
    filter(p_val_adj < 0.01) %>%
    arrange(avg_log2FC) %>%
    tail(n = 5) %>%
    select(gene) 
  top_genes <- top_genes[,1]
  
  if(length(rownames(top_genes))>0){
    
    for(gene in top_genes$gene){
      DefaultAssay(OHDA) <- "SCT"
      print(FeaturePlot(object = OHDA, features = gene, reduction = "umap_after_harmony", slot = "counts"))
    }
  }
}
```
